---
title: "Problem set 1: Quadratic equation"
author: "Frances Aponte"
#format: html
format: pdf
output:
  pdf_document:
    latex_engine: xelatex
editor: visual
---

## Excersise 1 
Defines variables $a=1, b=-1,c=-2$ and print out the solutions to $f(x)=ax^2+bx+c=0$.Do not report complex solutions, only real numbers. Avoid using the variable name c as it is a reserved function in R. Show the code and the answer.

```{r Define variables}
a=1
b=-1
# c will be called k
k=-2
```

```{r real solutions}
#To determine if the quadratic equation have real solutions we use the discriminant. 
real_soln<-b^2-4*a*k

if (real_soln>=0){
  soln_1<-(-b+sqrt(real_soln)/2*a)
  soln_2<-(-b-sqrt(real_soln)/2*a)
  
  cat("Real Solutions: \n")
  cat("x =", soln_1, "\n")
  cat("x =", soln_2, "\n")
} else {
  cat("No real solutions. \n")
}
```

## Excercise 2

Show a graph of $f(x)$ versus x for $x \in (-3,3)$. Do not show the code, only the graph.

```{r Graph, echo=FALSE}
library (ggplot2)
x<-seq(-3,3) #values for x
y<-a*x^2+b*x+k #values for y

ggplot(
  #values
  data.frame(x,y), aes(x,y)) +
  #defining the color for the quadratic function
  geom_line(color = "blue") + 
  #an horizontal line in the origin
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") + 
  #graph labeling
  labs(title = "Quadratic Equation Graph",
       x = "x",
       y = "f(x)")

```

## Excercise 3

Generate a PDF report.

To do it i modified the yaml to 

title: "Problem set 1: Quadratic equation"
author: "Frances Aponte"
format: pdf
output:
  pdf_document:
    latex_engine: xelatex
editor: visual







